---
title: "R Notebook"
output: html_notebook
---
```{r}
load("/Users/aizamalik/Desktop/BST260/Project/BST260_project_data.RDa")
```

```{r}
library(dplyr)
library(tidyverse)
library(maps)
library(stringr)
```

Purpose: to visualize which cuisines, grouped by country, are the most interacted with (have the most ratings) and which ones are most common (number of recipes of a specific cuisine).

```{r}
head(data_final)
world_map = map_data("world")
```

First, I created a data frame 'commonentries' that counts the instances of an entry and arranges them in descending order by frequency. Then I created a dataframe 'commonratings' that counts the number of ratings each cuisine entry type has and arranges them in descending order. I combined these two data frames into a new dataframe 'commonvalues2'. 

```{r}
#commonstrings <- data_final %>% add_count(frequency = cuisine) %>% arrange(., desc(n), by_group = frequency)

#By number of entries in data frame:
commonentries <- data_final %>% group_by(cuisine) %>% summarise(count=n()) %>% arrange(desc(count))

#By number of ratings
commonratings <- data_final %>% group_by(cuisine) %>% filter(!is.na(n_ratings)) %>% summarise(count=sum(n_ratings)) %>% arrange(desc(count))

#Ratings normalized for number of recipes
commonvalues <- left_join(commonentries, commonratings, by = "cuisine") 
commonvalues2 <- commonvalues %>% group_by(cuisine) %>% mutate(normalized = sum(count.y/count.x)) %>% arrange(desc(normalized))
```

To simplify the mapping analysis I chose to focus on entries that gave us country-specific information. I excluded any regional entries and created a new data frame with the top 10 entries from 'commonvalues2'. I created a new column to have cuisine types match up with regions in 'world map' and then combined these data frames. 

```{r}
common <- commonvalues2 %>% filter(cuisine %in% c("Mexican", "Italian", "Chinese", "Thai", "French", "Indian", "Greek", "Vietnamese", "American", "Asian, Japanese")) %>% arrange(desc(count.x))
Countries = c("Mexico", "Italy", "China", "Thailand", "France", "India", "Greece", "Vietnam", "United States", "Japan")
common$region <- Countries
```
```{r}
world_map_data <- left_join(world_map, common, by = "region")
```

World maps mapping the cuisines with the most number of recipes and which cuisine was interacted with most frequently (measured by number of ratings for each cuisine)

```{r}
library(RColorBrewer)
world_map_data %>% ggplot(aes(x = long, y = lat, group = group, fill = count.x)) +
   geom_polygon(color = "white") + scale_fill_gradientn(colors = brewer.pal(5, "Reds")) + 
   theme(panel.grid.major = element_blank(), 
         panel.background = element_blank(),
         axis.title = element_blank(), 
         axis.text = element_blank(),
         axis.ticks = element_blank()) + ggtitle("Most Popular Recipes")
```
```{r}
library(RColorBrewer)
world_map_data %>% ggplot(aes(x = long, y = lat, group = group, fill = count.y)) +
   geom_polygon(color = "white") + scale_fill_gradientn(colors = brewer.pal(5, "Reds")) + 
   theme(panel.grid.major = element_blank(), 
         panel.background = element_blank(),
         axis.title = element_blank(), 
         axis.text = element_blank(),
         axis.ticks = element_blank()) + ggtitle("Most Interacted-With Recipes")
```

